### 2016-12-02 real/draft.tex
draft.tex extensively rewritten
This rewrite accommodate the double limits for the same expansion of
p(Q|data). This draft is as yet incomplete and must still be finished in
Section 5.

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{161130}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}

@@

\item Traditional methods using conventional counting ratio
  (``relative frequency'') can also make predictions for such
  uncertainties, but only if the number of packets is large. By
  contrast, the Bayesian method presented here remains valid also for
  small packet multiplicities.

@@

We first sketch the generic notation and relations for Bayesian
probability theory.
%
A particular experiment or simulation generates
\textit{data}
which will often come in the form of a list of $n$ measurements,
$\bmy = (y_1,y_2,\ldots,y_n)$. In order to understand or explain this
data, the observer constructs one or more \textit{models} or
\textit{hypotheses} $\hypo$. Central to this construction is the
specification of a \textit{likelihood} $p(\bmy\cond \bmtheta,\hypo)$
where $\bmtheta = (\theta_1,\theta_2,\ldots,\theta_K)$ is the set of
unknown parameters specified by the model. The likelihood is the
probability assigned by the model $\hypo$ to the data $\bmy$ when the
parameters $\bmtheta$ are fixed and known.
%
If and when the individual measurements $y_i$ are independent, the
joint likelihood is the product
\begin{align}
  \lleq{bta}
  p(\bmy\cond \bmtheta) &= \prod_{i=1}^n p(y_i\cond \bmtheta).
\end{align}
At this level of analysis, the goal is to find which values for these
parameters provide a satisfactory description of the data. In the
Bayesian context, these parameter values are best described in terms
of the probability of the parameters given the data,
$p(\bmtheta\cond \bmy)$. This so-called \textit{posterior} is related
to the likelihood by Bayes' Theorem,
\begin{align}
  \lleq{bth}
  p(\bmtheta\cond \bmy)
  &= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    {p(\bmy)}
    %= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    %{\int d\bmtheta'\, p(\bmy\cond\bmtheta')\;p(\bmtheta')}
\end{align}
where $p(\bmtheta)$, the so-called \textit{prior}, is the probability
assigned to the values of the parameters before any data was taken
into account and the \textit{evidence}
\begin{align}
  \lleq{bti}
  p(\bmy) &= \int d\bmtheta'\, p(\bmy\cond\bmtheta')\;p(\bmtheta'),
\end{align}
is fully determined by the likelihood and prior.
%
Prediction of any ``future'' data $Y$ is made not by considering the
likelihood $p(Y\cond\bmtheta_{\rm max})$ using single best-fit
parameter values $\bmtheta_{\rm max}$, but by integrating over all
possible values of the $\bmtheta$ in terms of the posterior
$p(\bmtheta\cond n,\bml)$,
\begin{align}
  \lleq{btj}
  p(Y\cond \bmy) &= \int d\bmtheta\,p(Y\cond\bmtheta)\,p(\bmtheta\cond \bmy).
\end{align}
The above relations are true for any type of data and can be applied
to actual experimental data and the pertinent physical parameters. The
scope of this paper, however, is limited to the relationship between
simulated data generated by radiative transfer simulation codes.
Likewise, all parameters considered below relate strictly to model
descriptions of the simulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{From single-simulation data to many-simulation prediction}

% \subsection{Output data from radiative transfer simulations}
\subsection{Simulation data}

We consider a single computer simulation which generates a luminosity
spectrum from $\npack$ photon packets, each with a particular
frequency $\nu_i$ and luminosity $\lum_i$. Together, these frequencies
and luminosities form the raw simulation data
$\bmy = \{ (\nu_i,\lum_i): i = 1,\ldots,\npack\}$.  We divide up the
frequency spectrum into bins of width $\Delta\nu$ centered on
frequencies $\nu_b$ resulting in bin intervals
$\Delta_b = [\nu_b {-} \tfrac{1}{2} \Delta\nu \,,\, \nu_b {+}
\tfrac{1}{2} \Delta\nu], b = 1,2,\ldots, B$. Let $n_b$ be the number
of packets having a frequency falling into a particular bin $b$, and
let the vector $\bml_b = \{(\ell_{b,j}): j = 1,\ldots,n_b\}$ be the
set of luminosities of those packets. The data set is correspondingly
cast into the more compact histogram form
\begin{align}
  \lleq{sdb}
  \bmy = \{(n_b,\bml_b): b = 1,\ldots,B\}
\end{align}
with loss of information limited to replacing the exact packet
frequencies $\nu_i$ by the appropriate $\nu_b$. The total luminosity
in bin $b$ is the sum of luminosities of those packets whose frequency
falls into $\Delta_b$,
\begin{align}
  \lleq{sdc}
  \lumtot_b
  % &= \sum_{i=1}^\npack \bm{1}_{\Delta_b}(\nu_i)\,\ell_i
  &= \sum_{j=1}^{n_b} \lum_{b,j}.
\end{align}
% where the indicator function $\bm{1}_{[a,b]}(x) = 1$ when $x$ is in
% the interval $[a,b]$ and 0 otherwise.

\subsection{Prediction of luminosity spectrum}

Within a given model $\hypo$, the data $\bmy$ is only one of many
possible realisations of the corresponding random variables; for
example, starting the simulation with a different seed would result in
different output data. We denote the general random variables by the
corresponding upper-case symbols,
\begin{align}
  \bmL_b &= \{(\Lum_{b,j}): j = 1,\ldots,N_b\}\\
  \bmY &= \{(N_b, \bmL_b): b = 1,\ldots,B\},\\
  \Lumtot_b &= \sum_{j=1}^{N_b} \Lum_{b,j},
\end{align}
with corresponding likelihood $p(\bmY\cond\bmtheta)$ and posterior
$p(\bmtheta\cond\bmY)$ for any $\bmY$. The lower-case $(n_b,\bml_b)$
are strictly reserved for the simulated data at hand.

The goal of this paper is to illustrate how the single-simulation data
$\bmy$ can be used to predict the distribution of total bin
luminosities $Q_b, b = 1,\ldots,B$ if many simulations had been
carried out.

Various features such as absorption lines make the luminosity spectrum
highly variable. The model we develop below takes into account this
variability in a two-step process. First, the number of packets $N_b$
is modelled by means of a separate Poisson distribution
$p(N_b\cond\lambda_b)$ for each bin, where the parameter $\lambda_b$
will capture most of the variability of the spectrum.  Second, the
single-packet luminosity distribution is modelled by means of a packet
likelihood for $\Lum_{b,j}$ with packet parameters, a packet prior and,
of course, packet parameter posterior. The packet parameters, which
depend on the frequency bin, will be calculated in two ways: first by
drawing them from a ``parametrisation of the parameters'' across all
bins as shown in \refsec{example} and secondly by assuming that
\refsec{asymptotic}

Our approach
is to predict each bin's luminosity $Q_b$ separately using the same
model but allowing its parameters to vary independently in each bin.
Connections between them

function of frequency. In a binned analysis, we can avoid these
difficulties associated to modeling $n_b$. Our approach, however,
requires an explicit expression of how the distribution of the
luminosity of a single packet in our simulation changes with $\nu$. It
is this crucial piece of information that allows us to extend previous
methods in dealing with the extreme cases when the number of packets
in a bin, $n_b$, is one or even zero.

To keep it simple, we shall assume that all bins are independent,
meaning that
\begin{align}
  p(Q_1,Q_2,\ldots,Q_B\cond\bmy) = \prod_{b=1}^B p(Q_b\cond n_b,\bml_b).
\end{align}
Below, we calculate $p(Q_b\cond n_b,\bml_b)$ for fixed $b$ and hence
drop the subscript, i.e.\ $Q_b\to Q, n_b \to n, N_b\to N$ etc.

%%%%%%%%%%%%%%%%%%%
\fred{Refs: Zech, maybe others, too}

In terms of the generic notation of Section \ref{sec:bayp}, Bayesian
prediction of the distribution of the bin's luminosity
$Q = \sum_{j=1}^n L_j$ given the bin's data $(n,\bml)$ is obtained by
considering not only the likelihood $p(Q\cond\bmtheta_{\rm max})$
using best-fit parameter values $\bmtheta_{\rm max}$, but by
integrating over all possible values of the $\bmtheta$ in terms of the
posterior $p(\bmtheta\cond n,\bml)$,
\begin{align}
  p(Q\cond n,\bml) &= \int d\bmtheta\,p(Q\cond\bmtheta)\,p(\bmtheta\cond n,\bml).
\end{align}
The model's parameters $\bmtheta$ and calculation of this prediction
will be set out in detail below.

\section{Model} \label{sec:model}

We consider the total number of packets $\npack$ fixed, for example
chosen by a user. In contrast, the number of packets in a bin, $n_b$
is a realization of a random variable $N_b$ which we assume to follow
a Poisson distribution. With this assumption, we neglect bin-by-bin
correlations and treat each bin independently. From now on, we focus
on a single bin and drop the subscript $b$. Hence
\begin{align}
  \lleq{caa}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^N}{N!}
  \qquad N = 0,1,\ldots,\infty ,
\end{align}
where the dependency of the Poisson mean, $\lambda>0$, on the bin, or
respectively the frequency $\nu$, is implicit. Let $\bmL = (\Lum_1,\dotsc,\Lum_N)$ denote the
collection of luminosities of packets in the bin, then the total
luminosity in that bin is
% and its realization based on the data $\bmy$ are
\begin{align}
  \lleq{cab}
  \Lumtot &= \sum_{j=1}^{N} \Lum_j%, \qquad \lumtot = \sum_{j=1}^{n} \lum_j
  \,.
\end{align}
The distribution of the luminosity for each packet,
$p(\Lum \cond \bmphi)$, depends on parameters $\bmphi$ and potentially
on the frequency but we suppress the latter dependency.  The packets
are logically independent. Since $N$ follows a Poisson distribution,
$\Lumtot$ follows a \emph{compound Poisson process}.
%
The key quantity of interest is the distribution of $\Lumtot$ given
the data $\bmy$, $p(\Lumtot \cond \bmy)$ for the compound Poisson
process. In the following, we rewrite $p(\Lumtot \cond \bmy)$ to
obtain an expression that we can readily evaluate numerically. Using
the law of total probability,
\begin{align}
  \lleq{qaa}
  p(\Lumtot \cond \bmy) &= \int \rmdx{\bmphi} p(\bmphi \cond \bmy) p(\Lumtot \cond \bmphi, \bmy) \,,
\end{align}
where the first term in the integral is the posterior $p(\bmphi \cond \bmy)$ after observing $\bmy$. We can decompose the second term as
\begin{align}
  \lleq{qab}
  p(\Lumtot \cond \bmphi, \bmy) &= \sum_{N=0}^\infty  p(N \cond \bmphi, \bmy) p(\Lumtot \cond N, \bmphi, \bmy)\,,
\end{align}
where $p(N \cond \bmphi, \bmy) = p(N \cond \lambda)$ is given by
\refeq{caa} and the second term in the sum is
\begin{align}
  \lleq{qac}
  p(\Lumtot \cond N, \bmphi, \bmy) &= \int \rmdx{\bmL} p(\Lumtot \cond \bmL, N, \bmphi, \bmy) p(\bmL \cond N, \bmphi, \bmy)\\
  \lleq{qaf}
  &= \int \rmdx{\bmL} \delta \left( \Lumtot - \sum_{j=1}^N \Lum_j  \right) \prod_{j=1}^N p(\Lum_j \cond \bmphi)\,,
\end{align}
hence $Q$ does not depend on $\bmy$ explicitly. In the next
subsections, we provide more details on our assumptions and provide
explicit expressions for \refeq{qaa}--\refeq{qaf}.

% As the reference frequency, we choose the bin center.

\subsection{Modeling the packet luminosity} \label{sec:pack-lum}

In general, the distribution $p(\Lum_j \cond \bmphi)$ could be a of a
different functional form in each bin. For simplicity, we assume it is
identical but may depend on parameters $\bmphi$ that need to be
inferred from $\bmy$.  To simplify the treatment, we assume that the
frequency bins are narrow or equivalently that $p(\Lumtot \cond \bmy)$
does not vary within a bin. This implies that every sample $\Lum_j$ in
the bin follows the same distribution.

What is the role of $\bmphi$? It contains all parameters needed to
model the entire spectrum that have to be learned from $\bmy$. To
begin, we include the Poisson means for each bin
\begin{align}
  \lleq{qad} \bmphi = \left( \bmlambda, \bmxi \right) = \left(
    {\lambda_1,\dotsc,\lambda_{n_b}}, \bmxi \right)\,.
\end{align}
\fred{How }
The number and the luminosity of the packets are assumed independent,
so the posterior factorizes as
\begin{align}
  \lleq{qadd}
  p(\bmphi \cond \bmy) = p(\bmlambda \cond \bmy) p(\bmxi \cond \bmy) \,.
\end{align}

Regarding the additional parameters $\bmxi$, we have essentially two
extreme options; both illustrated below in \refsec{example} and
\refsec{asymptotic}. The first option is to assume an explicit
\emph{parametric} form, for example some function $f$, such that
\begin{align}
  \lleq{qae}
  p(\Lum \cond \bmphi) = f(\Lum \cond \bmxi, \nu) \,,
\end{align}
where we now made the dependency on $\nu$ explicit again.  In this
way, we assume the same distribution for all frequencies and we only
need to learn the probable values of $\bmxi$ from the data $\bmy$,
$p(\bmxi \cond \bmy)$. The advantage is that $f(\Lum \cond \bmxi, \nu)$ may
be well constrained by $\bmy$ if $\npack$ is large and it is easy to
include further assumptions; e.g., $f$ should be continuous in $\nu$.

The second option is preferable if it is impossible to elicit an
appropriate explicit expression for $p(\Lum \cond \bmphi)$ \emph{and}
each bin has many packets; i.e., $n \gg 1$. Then we can include one or
more parameters for each bin, consider each bin independently, and
perform a \emph{nonparametric} analysis. Compared to the parametric
approach, we now typically have many more parameters, and each one is
determined only by the packets in one instead of all bins.

\subsection{Poisson analysis}

In order to constrain the Poisson mean $\lambda$ in a bin, we have one
measurement, the number of packets $n$ in that bin. Using Bayes' theorem, we find the posterior
\begin{align}
  \lleq{pmd}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)} \,.
\end{align}
% to find the posterior predictive of $N$ given $n$ taking into account
% all possible values of $\lambda$,
% \begin{align}
%   \lleq{pme}
%   p(N\cond n)
%   &= \int \rmdx{\lambda}p(N,\lambda\cond n)
%   \ =\ \int \rmdx{\lambda} p(N\cond\lambda, n)\,p(\lambda\cond n)
%   \ =\ \frac{\int \rmdx{\lambda} p(N\cond\lambda)\,p(n\cond\lambda)\,p(\lambda)}
%   {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)}
% \end{align}
% where $p(N\cond\lambda,n) = p(N\cond\lambda)$ is independent of $n$.
Common uninformative priors are
\begin{align}
  \lleq{pmf}
  p(\lambda) = c \lambda^{-a},
\end{align}
where $a=0$ gives the uniform prior\footnote{For a uniform prior, the
  case $a{=}0$, we require of course a maximum $\lambda_{\rm max}$. As
  long as $\lambda_{\rm max}$ is large enough, the integral can
  effectively be taken to infinity.}, $a=1/2$ gives the Jeffreys and
reference prior, and $a=1$ is the transformation-group result
advocated by Jaynes~\cite[Ch. 12]{jaynes2003probability}. All choices
lead to an improper prior but the posterior is proper provided that
$n \ge 1$; i.e., at least one packet is observed in the bin. The
standard result for the posterior is
\begin{align}
  \lleq{pms}
  p(\lambda\cond n) &= \GammaDist(\lambda \cond \alpha = n-a+1, \beta = 1)\,,
\end{align}
where the  $\GammaDist$ distribution with shape parameter $\alpha$ and rate
parameter $\beta$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(\lambda | \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}\, , \; x > 0 .
\end{align}
% \begin{align}
%   \lleq{pmg}
%   p(n) &= %
%   \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda)
%   \ =\
%   c \int_0^\infty d\lambda\, \frac{e^{-\lambda}\,\lambda^{n-a}}{n!}
%   \ =\ c\frac{(n-a)!}{n!} \, ,
%   \\
%   \lleq{pmh}
%   p(N\cond n) %
%   &= \frac{n!}{(n-a)!} \int_0^\infty d\lambda\,
%   \frac{e^{-2\lambda} \lambda^{N+n-a}}{N!\, n!}
%   \ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
% \end{align}

\section{Example application: \tardis} \label{sec:example}

 The \tardis{} packages simulates supernovae with \fred{Wolfgang's
  favorite description of his package}. Here we focus on the real
packets as opposed to the virtual packets output by tardis.

>>>>

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{161202}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}

@@

\item Traditional methods using conventional counting ratios
  (``relative frequencies'') can also make predictions for such
  uncertainties, but only if the number of packets is large. By
  contrast, the Bayesian method presented here remains valid also for
  small packet multiplicities.

@@

We first sketch the generic notation and relations for Bayesian
probability theory.
%
A particular experiment or simulation generates
\textit{data}
which will often come in the form of a list of $n$ measurements,
$\bmy = (y_1,y_2,\ldots,y_n)$. In order to understand or explain this
data, the observer constructs one or more \textit{models} or
\textit{hypotheses} $\hypo$. Central to this construction is the
specification of a \textit{likelihood} $p(\bmy\cond \bmtheta,\hypo)$
where $\bmtheta = (\theta_1,\theta_2,\ldots,\theta_K)$ is the set of
unknown parameters specified by the model. The likelihood is the
probability assigned by the model $\hypo$ to the data $\bmy$ when the
parameters $\bmtheta$ are fixed and known.
%
If and when the individual measurements $y_i$ are independent, the
joint likelihood is the product
\begin{align}
  \lleq{bta}
  p(\bmy\cond \bmtheta) &= \prod_{i=1}^n p(y_i\cond \bmtheta).
\end{align}
At this level of analysis, the goal is to find which values for these
parameters provide a satisfactory description of the data. In the
Bayesian context, these parameter values are best described in terms
of the probability of the parameters given the data,
$p(\bmtheta\cond \bmy)$. This so-called \textit{posterior} is related
to the likelihood by Bayes' Theorem,
\begin{align}
  \lleq{bth}
  p(\bmtheta\cond \bmy)
  &= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    {p(\bmy)}
    %= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    %{\int d\bmtheta'\, p(\bmy\cond\bmtheta')\;p(\bmtheta')}
\end{align}
where $p(\bmtheta)$, the so-called \textit{prior}, is the probability
assigned to the values of the parameters before any data was taken
into account and the \textit{evidence}
\begin{align}
  \lleq{bti}
  p(\bmy) &= \int d\bmtheta'\, p(\bmy\cond\bmtheta')\;p(\bmtheta'),
\end{align}
is fully determined by the likelihood and prior.
%
Prediction of any ``future'' data $Y$ is made not by considering the
likelihood $p(Y\cond\bmtheta_{\rm max})$ using single
maximum-likelihood parameter values $\bmtheta_{\rm max}$, but by
integrating over all possible values of the likelihood, weighted by
the posterior,
\begin{align}
  \lleq{btj}
  p(Y\cond \bmy) &= \int d\bmtheta\,p(Y\cond\bmtheta)\,p(\bmtheta\cond \bmy).
\end{align}
The above relations are true for any type of data and can be applied
to actual experimental data and the pertinent physical parameters. The
scope of this paper, however, is limited to the relationship between
simulated data generated by radiative transfer simulation codes.
Likewise, all parameters considered below relate strictly to model
descriptions of the simulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Single-simulation data and multi-simulation prediction}

% \subsection{Output data from radiative transfer simulations}
% \subsection{Simulation data}

We consider a single computer simulation which generates a luminosity
spectrum from $\npack$ photon packets, each having a particular
frequency $\nu_i$ and luminosity $\lum_i$. Together, these frequencies
and luminosities form the raw simulation data
$\bmy = \{ (\nu_i,\lum_i): i = 1,\ldots,\npack\}$.  We divide up the
frequency spectrum into bins of width $\Delta\nu$ centered on
frequencies $\nu_b$ resulting in bin intervals
$\Delta_b = [\nu_b {-} \tfrac{1}{2} \Delta\nu \,,\, \nu_b {+}
\tfrac{1}{2} \Delta\nu], b = 1,2,\ldots, B$. Let $n_b$ be the number
of packets having a frequency falling into a particular bin $b$, and
let the vector $\bml_b = \{(\ell_{b,j}): j = 1,\ldots,n_b\}$ be the
set of luminosities of those packets. The data set is correspondingly
cast into the more compact histogram form
\begin{align}
  \lleq{sdb}
  \bmy = \{(n_b,\bml_b): b = 1,\ldots,B\}
\end{align}
with loss of information limited to replacing the exact packet
frequencies $\nu_i$ by the approximate bin frequency $\nu_b$. The
total luminosity in bin $b$ is
\begin{align}
  \lleq{sdc}
  \lumtot_b 
  &= \sum_{j=1}^{n_b} \lum_{b,j}.
\end{align}
Within a given model $\hypo$, the data $\bmy$ is only one of many
possible realisations of the corresponding random variables; for
example, starting the simulation with a different seed would result in
different output data. We denote the general random variables by the
corresponding upper-case symbols,
\begin{align}
  \lleq{sdd}
  \bmL_b &= \{(\Lum_{b,j}): j = 1,\ldots,N_b\}\\
  \lleq{sde}
  \bmY &= \{(N_b, \bmL_b): b = 1,\ldots,B\},\\
  \lleq{sdf}
  \Lumtot_b &= \sum_{j=1}^{N_b} \Lum_{b,j},
\end{align}
Lower-case symbols such as $\bmy, n_b,\bml_b$ etc.\ are strictly
reserved for the simulated data at hand.

The goal of this paper is to illustrate how the single-simulation data
$\bmy$ can be used to predict the distribution of total bin
luminosities $Q_b, b = 1,\ldots,B$ if many simulations had been
carried out. At this introductory level, we assume that the different
luminosities are independent, so that the prediction of the entire
spectrum (including all its uncertainty) is the product of predictions
for each frequency bin,
\begin{align}
  \lleq{sdg}
  p(Q_1,Q_2,\ldots,Q_B\cond\bmy) = \prod_{b=1}^B p(Q_b\cond \bmy).
\end{align}

\section{The generic model} \label{sec:model}

Various physical processes such as absorption result in luminosity
spectra with strong features. The model we develop below takes into
account the variability introduced by these features in a two-step
process termed a \textit{compound process} in the statistics
literature. 
% 
In the first step, the number of packets $N$ is modelled by means of a
separate Poisson distribution $p(N\cond\lambda)$ for each bin, where
the magnitude of $\lambda$ will capture most of the variability of the
spectrum. Given $N$, the single-packet luminosity distribution is
modelled in the second step by means of a packet likelihood for
$\Lum_{b,j}$ with packet parameters, a packet prior and, of course,
packet parameter posterior. The single-packet distribution contains
the crucial information which allows us to extend previous methods to
cases where the number of packets in a bin is small.
%%%%%%%%%%%%%%%%%%%
\fred{Refs: Zech, maybe others, too}

We derive in this section the generic equations of the compound
process, and then proceed to show in Sections \ref{sec:example} and
\ref{sec:asymptotic} how these equations can be used in two different
ways.
% 
We consider for the moment quantities in one bin $b$ only, and hence
drop the $b$-subscripts, e.g.\
$Q_b \to Q, n_b \to n, N_b\to N, \bml_b\to\bml, L_{b,j} \to L_j$ etc.

We seek a Bayesian prediction for the total luminosity in one bin,
$p(Q\cond n,\bml)$, one of the factors in Eq.~(\ref{sdg}).  To do so,
we must take into account all possible values of variables and
parameters entering our model.  This involves multiple use of the
\textit{marginalisation rule}, which states that, to obtain the
probability for a variable $U$ when we are not interested in another
variable $V$, the marginal probability for $U$ is %
$p(U) = \int dV\,p(U,V) = \int dV\,p(U\cond V)\,p(V)$ for continuous
$V$ and $p(U) = \sum_V p(U,V) = \sum_V p(U\cond V)\,p(V)$ for discrete
$V$ respectively.
% 
The uninteresting variables $V$ in the present case are the number of
packets $N$, the set of luminosities $\bmL$ of the $N$ packets
whose likelihood is modelled by a distribution determined by
likelihood parameters $\bmphi$,
\begin{align}
  \lleq{slk}
  p(\bmL\cond N,\bmphi)
  &= \prod_{j=1}^{N} p(L_j\cond \bmphi)
\end{align}
and the parameter $\lambda$ governing the likelihood for the number of
packets, $p(N\cond\lambda)$. Multiple use of the marginalisation rule
on $p(Q\cond n,\bml)$ results in
\begin{align}
  \lleq{sll}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;\int_0^\infty d\lambda\;
    p(N\cond\lambda)\;
    p(\lambda\cond n)
    \int d\bmL\;
    p(Q\cond N,\bmL)\;
    \int d\bmphi\;
    p(\bmL\cond N,\bmphi)\,
    p(\bmphi\cond n,\bml),
\end{align}
where we have omitted from the probabilities all quantities to the
right of the conditional line which are irrelevant to the variable at
hand. For example, Eq.~(\ref{sdf}) shows that the total bin luminosity
$Q$ is fully determined by $N$ and $\bmL$ and so we can shorten
$p(Q\cond N,\bmL,\lambda,\bmphi,n,\bml)$ to $p(Q\cond N,\bmL)$. %
Likewise the probability for $\bmL$ is fully determined by $N$ and the
parameters $\bmphi$ and does not depend on the data $n,\bml$.  The
bounds of $\int d\bmL$ and $\int d\bmphi$ are determined by the
specific model and data for which $p(Q\cond n,\bml)$ is to be
calculated.
% 
We note that a part of Eq.~(\ref{sll}) can be seen as a convolution
since $p(Q\cond N,\bmL)$ can be written as a Dirac delta function,
\begin{align}
  \lleq{ddf}
  \int d\bmL\; p(Q\cond N,\bmL)\; p(\bmL\cond N,\bmphi)
  &= \int d\bmL\;\delta(Q - \textstyle\sum_{j=1}^N L_j)\; \prod_j p(L_j\cond \bmphi).
\end{align}
Eq.~(\ref{sll}) is exact, and in principle all that remains is to make
specific choices for the various probabilities which are judged best
to represent the data at hand, and thence to carry out the sum and
multiple integrals. In practice, of course, such calculations may be
time-consuming, and there will often be more than one choice for the
functional form of each probability. In Section \ref{sec:example}, we
shall therefore set out a specific example

\section{Example A: modelling single-packet luminosity distributions} 
\label{sec:example}

In this section, we show by example how information on the
single-packet luminosity distributions can be utilised to find exact
expressions for Eq.~(\ref{sll}).

\subsection{Multiplicity prediction}

First, the parameter $\lambda$ is eliminated as follows. The number of
packets $N$ is Poisson-distributed
\begin{align}
  \lleq{caa}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^{N}}{N!}
  \quad N = 0,1,\ldots,\infty.
\end{align}
Data $n$ is naturally presumed to have arisen from the same Poisson
distribution with the same $\lambda$, so $p(n\cond\lambda) =
e^{-\lambda}\lambda^{n}/n!$. %
% and since $n$ is just one realisation of $N$, it likewise is
% poissonian $p(n\cond\lambda) = e^{-\lambda}\lambda^{n}/n!$ with the
% same $\lambda$.
The posterior for $\lambda$ is found from the data $n$ via Bayes'
Theorem,
\begin{align}
  \lleq{pmd}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)} \,.
\end{align}
We use an improper prior 
\begin{align}
  \lleq{pmf}
  p(\lambda) = c \lambda^{-a},
\end{align}
meaning that $\int_0^\infty d\lambda\,p(\lambda)$ is known only up to
a constant $c$. A hyperparameter choice $a=0$ yields the uniform
prior, $a=1/2$ gives the Jeffreys and reference prior, and $a=1$ is
the transformation-group result advocated by
Jaynes~\cite[Ch. 12]{jaynes2003probability}. An improper prior is not
a problem because $c$ cancels out when calculating the posterior,
provided that $n \ge 1$; i.e., at least one packet is observed in the
bin.\footnote{For a uniform prior, the case $a{=}0$, we require of
  course a parameter $\lambda_{\rm max}$. As long as
  $\lambda_{\rm max}$ is large enough, it too would cancel.}, The
posterior is
\begin{align}
  \lleq{pms}
  p(\lambda\cond n) 
  &= \GammaDist(\lambda \cond \alpha = n{-}a{+}1 \,,\, \beta = 1)\,,
\end{align}
where the $\GammaDist$ distribution with shape parameter $\alpha > 0$ and
rate parameter $\beta > 0$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(x\cond \alpha, \beta) 
  &= \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}\,,
  \quad 0 \leq x < \infty.
\end{align}
The evidence is 
$p(n) =  \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda) = c\,(n-a)!/n!$
% \begin{align}
%   \lleq{pmg}
%   p(n) 
%   &=  \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda)
%     \ =\ c\frac{(n-a)!}{n!} \,, \\
% \end{align}
and inserting Eqs.~(\ref{caa}) and (\ref{pmd}), the prediction for $N$
given $n$ is
\begin{align}
  \lleq{pmh}
  p(N\cond n) %
  &= \int_0^\infty d\lambda\, p(N\cond\lambda)\,p(\lambda\cond n)
  \ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
\end{align}
so that Eq.~(\ref{sll}) simplifies to
\begin{align}
  \lleq{gmh}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int d\bmL\;
    p(Q\cond N,\bmL)\;
    \int d\bmphi\;
    p(\bmL\cond N,\bmphi)\,
    p(\bmphi\cond n,\bml),
\end{align}

\subsection{Single-packet description}

Next, we consider detailed modelling of $p(\bmL\cond N,\bmphi)$ which
was written in Eq.~(\ref{slk}) as a product single-packet
distributions $p(L_j\cond \bmphi)$, meaning that single packets are
assumed to be mutually independent and to follow the same distribution
independent of $N$.

Which specific probability function to choose for $p(L\cond\bmtheta)$
depends, of course, on the specific simulation data. In this section,
we choose it to be the same Gamma distribution for all $L_j$,
with parameters $\bmphi = (\alpha,\beta)$,
\begin{align}
  \lleq{spc}
  p(L_j\cond \bmphi) &\sim \GammaDist(L_j\cond \alpha,\beta)
\end{align}
which has the nice property that the convolution (\ref{ddf}) results
in yet another Gamma distribution,
\begin{align}
  \lleq{spd}
  \int d\bmL\; p(Q\cond N,\bmL)\; p(\bmL\cond N,\bmphi)
  &= \GammaDist(Q\cond N\alpha, \beta).
\end{align}
and so (\ref{gmh}) becomes in this example
\begin{align}
  \lleq{spe}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int d\alpha\,d\beta\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    p(\alpha,\beta\cond n,\bml).
\end{align}
The posterior for the parameters is calculated by yet another 
application of Bayes' Theorem,
\begin{align}
  \lleq{spf}
  p(\alpha,\beta\cond n,\bml)
  &= \frac{p(n,\bml\cond \alpha,\beta)\,p(\alpha,\beta)}{p(n,\bml)}
\end{align}
where the data $(n,\bml)$ must naturally follow the same
Gamma-distribution likelihood as $(N,\bmL)$, i.e.\
\begin{align}
  \lleq{spg}
  p(n,\bml\cond \alpha,\beta) 
  & = \prod_{j=1}^n \GammaDist(\ell_j\cond\alpha,\beta)
    \ =\ \frac{\beta^{n\alpha} p^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n}
\end{align}
with compact representation of the data in terms of
$q = \sum_{j=1}^n \ell_j$ and $r = \prod_{j=1}^n \ell_j$.
